---
title: "getting-started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(imhea)
```

```{r}
iMHEA_Catchment_AREA = read_csv(
  system.file("extdata", "iMHEA_Data_Areas.csv", package = "imhea"),
  show_col_types = FALSE
)
```

```{r}
q1_raw <- read_csv(
  system.file("extdata", "LLO/iMHEA_LLO_01_HI_01_raw.csv", package = "imhea"),
  show_col_types = FALSE
)

p1_raw = read_csv(
  system.file("extdata", "LLO/iMHEA_LLO_01_PO_01_raw.csv", package = "imhea"),
  show_col_types = FALSE
)

p2_raw = read_csv(
  system.file("extdata", "LLO/iMHEA_LLO_01_PO_02_raw.csv", package = "imhea"),
  show_col_types = FALSE
)
```

Convert precipitation data to rain_gauge objects:

```{r}
p1 <- p1_raw %>% tipping_bucket_rain_gauge(id = "LLO_01_P0_01", event_units = "mm")
p2 <- p2_raw %>% tipping_bucket_rain_gauge(id = "LLO_01_P0_02", event_units = "mm")
```

Convert streamflow data to stream_gauge object:

```{r}
q1 <-
  q1_raw %>%
  stream_gauge(
    id = "LLO_01_HI_01",
    discharge_units = "l/s",
    level_column = "Level cm",
    level_units = "cm"
  )
```

Define the catchment ID, and retrieve catchment area:
```{r}
catchment_id <- "LLO_01"
catchment_area <-
  iMHEA_Catchment_AREA %>%
  filter(Catchment %in% catchment_id) %>%
  `[`(, 2, drop=T)
```

Create a catchment object (this takes a few minutes):

```{r}
x <- catchment(q1, p1, p2, id = catchment_id, area = set_units(catchment_area, km^2))
```

Have a look at the catchment indices:

```{r}
indices(x)
```

Make some plots:

```{r}
p1 <- ggplot(data = daily(x), aes(x = Date, y = P)) +
  geom_line()

p2 <- ggplot(data = daily(x), aes(x = Date, y = Q)) +
  geom_line()
```

Join plots using `patchwork`

```{r}
library(patchwork)
p1 + p2 + plot_layout(ncol = 1, nrow = 2)
```

Plot monthly/annual summaries

```{r}
p1 <- ggplot(data = monthly(x), aes(x = Date, y = Q)) +
  geom_bar(stat = "identity")

p2 <- ggplot(data = annual(x), aes(x = Date, y = Q)) +
  geom_bar(stat = "identity")

p1 + p2 + plot_layout(ncol = 1, nrow = 2)
```

Plot IDC/FDC

```{r}
p1 <- ggplot(data = idc(x), aes(x = as.numeric(D), y = Intensity)) +
  geom_line() +
  scale_x_continuous(trans = "log10") +
  xlab("Duration [min]")

p2 <- ggplot(data = fdc(x), aes(x = Exceedance_Pct, y = Q)) +
  geom_line() +
  scale_y_continuous(trans = "log10") +
  ylab("Discharge [m3/s]")
```

Plot baseflow

```{r}
d <- baseflow_data(x)[["UKIH"]]
d <- d %>% pivot_longer(cols = starts_with("Q"), names_to = "name", values_to = "values")
p1 <- ggplot(data = d, aes(x = Date, y = values, col = name)) +
  geom_line()
```

Monthly climatology with box & whisker

```{r}
d <- monthly(x) %>% mutate(Month = month(Date))
p1 <- ggplot(data = d, aes(x = factor(Month), y = Q)) +
  geom_boxplot()

p2 <- ggplot(data = d, aes(x = factor(Month), y = P)) +
  geom_boxplot()

p1 + p2 + plot_layout(ncol = 1, nrow = 2)
```
